Draft_Application_Stochastic_Volatility

### 2021-07-16 Real Dataset SV.Rmd
layout

Stochastic Volatility Models, whose early formulation is commonly attributed to Taylor (1982, 1986), allow to account for time-varying and autoregressive volatility in financial returns, posing themselves as a valid alternative to ARCH (Engle 1982) or GARCH (Bollerslev 1986) models in dealing with non-constant volatility. Kim, Shephard and Chib (1998) define the canonical model for regularly spaced data as \begin{align*}

@@

\end{align*} where the log volatility $x_t$ is assumed to follow a stationary process ($|\beta|<1$), $\psi$ is a constant scaling factor, $\beta$ is the persistence in the volatility and $\tau$ is the volatility of the log-volatility.\footnote{For identifiability reasons, either $\psi=1$ or $\mu=0$. Kim, Shephard and Chib (1998) prefer $\psi=1$.} Most importantly, $\epsilon_t$ and $\eta_t$, the Gaussian white noise processes that drive the canonical model, are assumed to be uncorrelated. This latter assumption justifies another definition of this specification, namely "discrete SV model without leverage".\footnote{Note that in this review we do not focus on continuous time SV models. In fact, such models have attracted a considerable amount of research in financial econometrics and mathematical finance, especially after Hull and White (1987) considered stochastic volatility for option pricing. Arguably, the most influential model was then proposed for option pricing by Heston (1993), a SV model with leverage effects and square root diffusion driving volatility. Diffusion-based SV models enjoyed increasing popularity, see for example Barndorff-Nielsen and Shephard (2001), or Christoffersen et al. (2010), who investigate alternatives to the entrenched affine square root SV model.  Eraker (2004) proposed a SV model with correlated jumps in prices and volatility, extending Heston's model, while Comte and Renault (1998) extended Hull and White's model as to feature long memory properties. More recently, Gatheral et al. (2018) further built on Comte and Renault's fractional SV model to propose the popular "rough volatility models" (see also Friz et al., 2021).}\

@@

\end{align*} where $\rho<0$ captures the negative correlation. Such specification, which can be referred to as "discrete SV model with leverage", captures the asymmetric response of volatility to returns of different signs, so that similar specifications are sometimes also deemed as "asymmetric SV models" (e.g. Harvey and Shephard 1996, Mao et al. 2020).\

@@

Assuming that $\epsilon_t$ follows a Student-$t$ distribution, and exploiting the fact that $\epsilon_t$ can then be written as $\lambda_t^{-1/2}\zeta_t$, where $\zeta_t \sim N(0,1)$ and $v\lambda_t \sim \chi^2_v$ (Harvey et al. 1994, Chib et al. 2002), we have the following SV model with both fat tails and leverage effect (Jacquier et al., 2004), \begin{align*}

@@

\end{align*} To capture other elements of the behaviour of financial data, several other extensions have been proposed. For example, SV models have been extended to include conditional heteroskedasticity in the mean returns (Koopman and Hol Uspensky 2002) to capture potential volatility feedback effects, or to feature autoregressive moving average innovations (Chan 2013, Zhang et al. 2020), allowing better goodness of fit and out-of-sample forecasts.\footnote{Dimitrakopoulos and Kolossiatis (2020) note that "the moving average component, the leverage effect and the conditional heteroscedasticity in mean have been considered separately in the stochastic volatility literature" and provide two specifications, one featuring an MA component and leverage effects, the other an MA component and conditional heteroskedasticity in mean.} \footnote{Other extensions, though ones for which we do not report the specifications, feature modelling the latent volatility process $x_t$ as an ARFIMA process (Long Memory Stochastic Volatility model, Breidt et al. 1998) or as governed by a first-order Markov process (Markov Switching Stochastic Volatility model, So et al. 1998). Recently, Luo et al. (2018) incorporated neural networks in the stochastic volatility model (Neural Stochastic Volatility Model), while Xu and Chen (2021) employ deep learning models (Deep Stochastic Volatility Model).} The model proposed by Koopman and Hol Uspensky is specified as \begin{align*}

@@

\end{align*} while the state space representation of the ARMA(p,q)-SV framework, as in Zhang et al. (2020), reads \begin{align*}

>>>>

Stochastic Volatility Models, whose early formulation is commonly attributed to Taylor (1982, 1986), allow to account for time-varying and autoregressive volatility in financial returns, posing themselves as a valid alternative to ARCH (Engle 1982) or GARCH (Bollerslev 1986) models in dealing with non-constant volatility. Kim, Shephard and Chib (1998) define the canonical model for regularly spaced data as 
\begin{align*}

@@

\end{align*} 
where the log volatility $x_t$ is assumed to follow a stationary process ($|\beta|<1$), $\psi$ is a constant scaling factor, $\beta$ is the persistence in the volatility and $\tau$ is the volatility of the log-volatility.\footnote{For identifiability reasons, either $\psi=1$ or $\mu=0$. Kim, Shephard and Chib (1998) prefer $\psi=1$.} Most importantly, $\epsilon_t$ and $\eta_t$, the Gaussian white noise processes that drive the canonical model, are assumed to be uncorrelated. This latter assumption justifies another definition of this specification, namely "discrete SV model without leverage".\footnote{Note that in this review we do not focus on continuous time SV models. In fact, such models have attracted a considerable amount of research in financial econometrics and mathematical finance, especially after Hull and White (1987) considered stochastic volatility for option pricing. Arguably, the most influential model was then proposed for option pricing by Heston (1993), a SV model with leverage effects and square root diffusion driving volatility. Diffusion-based SV models enjoyed increasing popularity, see for example Barndorff-Nielsen and Shephard (2001), or Christoffersen et al. (2010), who investigate alternatives to the entrenched affine square root SV model.  Eraker (2004) proposed a SV model with correlated jumps in prices and volatility, extending Heston's model, while Comte and Renault (1998) extended Hull and White's model as to feature long memory properties. More recently, Gatheral et al. (2018) further built on Comte and Renault's fractional SV model to propose the popular "rough volatility models" (see also Friz et al., 2021).}\

@@

\end{align*} 
where $\rho<0$ captures the negative correlation. Such specification, which can be referred to as "discrete SV model with leverage", captures the asymmetric response of volatility to returns of different signs, so that similar specifications are sometimes also deemed as "asymmetric SV models" (e.g. Harvey and Shephard 1996, Mao et al. 2020).\

@@

Assuming that $\epsilon_t$ follows a Student-$t$ distribution, and exploiting the fact that $\epsilon_t$ can then be written as $\lambda_t^{-1/2}\zeta_t$, where $\zeta_t \sim N(0,1)$ and $v\lambda_t \sim \chi^2_v$ (Harvey et al. 1994, Chib et al. 2002), we have the following SV model with both fat tails and leverage effect (Jacquier et al., 2004), 
\begin{align*}

@@

\end{align*} 
To capture other elements of the behaviour of financial data, several other extensions have been proposed. For example, SV models have been extended to include conditional heteroskedasticity in the mean returns (Koopman and Hol Uspensky 2002) to capture potential volatility feedback effects, or to feature autoregressive moving average innovations (Chan 2013, Zhang et al. 2020), allowing better goodness of fit and out-of-sample forecasts.\footnote{Dimitrakopoulos and Kolossiatis (2020) note that "the moving average component, the leverage effect and the conditional heteroscedasticity in mean have been considered separately in the stochastic volatility literature" and provide two specifications, one featuring an MA component and leverage effects, the other an MA component and conditional heteroskedasticity in mean.} \footnote{Other extensions, though ones for which we do not report the specifications, feature modelling the latent volatility process $x_t$ as an ARFIMA process (Long Memory Stochastic Volatility model, Breidt et al. 1998) or as governed by a first-order Markov process (Markov Switching Stochastic Volatility model, So et al. 1998). Recently, Luo et al. (2018) incorporated neural networks in the stochastic volatility model (Neural Stochastic Volatility Model), while Xu and Chen (2021) employ deep learning models (Deep Stochastic Volatility Model).} The model proposed by Koopman and Hol Uspensky is specified as 
\begin{align*}

@@

\end{align*} 
while the state space representation of the ARMA(p,q)-SV framework, as in Zhang et al. (2020), reads 
\begin{align*}
